---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02

# Initial Plans

I was initially planning to explore the Florida Lakes spatial dataset more. I wanted to highlight the counties with the most area of lakes. I found it hard to try to merge the geoms and find the intersections to get this to work. I also wanted to try to show the correlation of house prices to certain futures of houses, I figured scatter plots would be the best way to visualize this.

# Narrative

We load in all of the different packages and data for our analysis in this first chunk. Sadley nothing too fun here…

```{r}
library(tidyverse)
library(sf)
library("rnaturalearth")
library("rnaturalearthdata")
library(plotly)
library(broom)
library(units)
library("tools")
library("maps")
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
counties <- st_as_sf(map("county", plot = FALSE, fill = TRUE))


setwd( "..")
florida_lakes <- read_sf("data/Florida_Lakes/Florida_Lakes.shp")

houses <- read_csv("data/WestRoxbury.csv")

```

Next, we can begin to take a look at the houses dataset. I found it very interesting that the largest average house sizes per year were in the 1800s. I would like to dive more into this in the future.

```{r}
houses %>%
  group_by(`YR BUILT`) %>%
  summarise(Average_Size = mean(`LIVING AREA`, na.rm = TRUE),
            Average_Value = min(`TOTAL VALUE`, na.rm = TRUE),
            Average_Users_Rooms = max(ROOMS, na.rm = TRUE)) %>%
  arrange(desc(Average_Size))
```

Below is an interactive plot that you can use to look at the houses dataset comparing the total value of a house to its living area. Feel free to explore it and take a look at some of the outliers.

```{r}

plot <- ggplot(houses, aes(x = `LIVING AREA`, y = `TOTAL VALUE`, size = BEDROOMS, color = BEDROOMS)) +
  geom_point(alpha = 0.7) +
  labs(title = "Total Value vs Living Area",
       x = "Living Area (sqft)",
       y = "Total Value (in $1000s)",
       size = "Number of Bedrooms",
       color = "Number of Bedrooms") +
  scale_size_continuous(range = c(.1, 3)) +
  theme_minimal()

# Convert ggplot2 to ggplotly for interactivity
inter_plot <- ggplotly(plot)

inter_plot

htmlwidgets::saveWidget(inter_plot, "value_vs_area.html")

```

Next we move to the Florida lakes data, we can see how lakes are spread through different counties. I limited this to central Florida so we can see lakes we are familiar with and pointed out the largest one.

```{r}
invalid_geometries <- !st_is_valid(florida_lakes)

florida_lakes <- st_make_valid(florida_lakes)

florida_lakes <- florida_lakes %>%
  mutate(area = st_area(geometry)) %>%
  arrange(desc(area))

florida_lakes$area <- as.numeric(florida_lakes$area)

lake_summary <- florida_lakes %>%
  group_by(COUNTY) %>%
  summarise(Average_Area = mean(area, na.rm = TRUE),
    total_lakes = n(),
    total_area = sum(area, na.rm = TRUE),
    mean_area = mean(area, na.rm = TRUE),
    median_area = median(area, na.rm = TRUE),
    max_area = max(area, na.rm = TRUE),
    min_area = min(area, na.rm = TRUE)) %>%
  arrange(desc(Average_Area))

lake_summary

```


```{r}
largest_lake <- florida_lakes[1, ]
centroid <- st_centroid(largest_lake$geometry)
label <- largest_lake$NAME

x_shift <- -.7
y_shift <- -.3

ggplot(data = florida_lakes) +
  geom_sf(data = states) +
  geom_sf(data = counties) +
  geom_sf(aes(fill = area)) +
  scale_fill_viridis_c(option = "plasma") +
  coord_sf(xlim = c(-83, -80), ylim = c(26.5, 29), expand = FALSE) +
  ggtitle("Map of Central Florida Lakes by Area") +
  theme_minimal() +
  labs(fill = "Area (sq units)") +
  annotate("text", x = st_coordinates(centroid)[, "X"] + x_shift , 
             y = st_coordinates(centroid)[, "Y"] + y_shift, 
             label = label, color = "red", size = 5, hjust = 0)

```
Below is the houses dataset again mapped with a linear model correlation. Much to our expectation we can visualize that house size has large correlation to the price. 

```{r}

ggplot(houses, aes(x = `TOTAL VALUE`, y = `LIVING AREA`)) +
  geom_point() +
  geom_smooth(method = "lm",
              formula = "y ~ x") +
  theme_minimal()

```

This is then backed up by the below correlation coefficients.

```{r}
house_model <- lm(`LIVING AREA` ~ `TOTAL VALUE`, houses)
houses_coefs <- tidy(house_model, conf.int = TRUE) %>% 
  filter(term != "(Intercept)")
houses_coefs
```
Overall, this project should highlight some of the basic ways to look at data. While this is just the surface these visualizations could be improved upon to add in more factors. I found the spatial data difficult to work with and would like to dive into this more. I find it to be a very interesting topic as well.